---
title: "czarnika_stats415_hw5"
author: "Alex Czarnik"
date: "February 23, 2016"
output: pdf_document
---

### 1a) Fit a logistic regression model that predicts Direction using Lag1 and Lag2.
```{r}
library(MASS)
library(ISLR)
attach(Weekly)
week = Weekly
glm.fit = glm(Direction ~ Lag1 + Lag2, data=week, family=binomial)
summary(glm.fit)
```

### 1b) Fit a logistic regression model that predicts Direction using Lag1 and Lag2 using all but the first observation.
```{r}
sub = week [-(1:1),]
glm.fit = glm(Direction ~ Lag1 + Lag2, data=sub, family=binomial)
summary(glm.fit)
```

### 1c) Use the model from (b) to predict the direction of the first observation. You can do this by predicting that the first observation will go up if Pr(Direction=“Up” | Lag1, Lag2) > 0.5. Was this observation correctly classified?
```{r}
first = week[1:1,]
glm.probs=predict(glm.fit, first, type="response")
glm.pred=rep("Down",1)
glm.pred[glm.probs>.5]="Up"
sprintf("glm.pred: %s, first$Direction: %s", glm.pred, first$Direction)
```
This observation was not correctly classified.

### 1d) Write a for loop from i = 1 to i = n, where n is the number of observations in the data set, that performs each of the following steps:

#### 1d i)  Fit a logistic regression model using all but the ith observation to predict Direction using Lag1 and Lag2.
#### 1d ii)   Compute the posterior probability of the market moving up for the ith observation
#### 1d iii) Use the posterior probability for the ith observation in order to predict whether or to predict whether or not the market moves up. 
#### 1d iv) Determine whether or not an error was made in predicting the direction for the ith observation. If an error was made, then indicate this as a 1, and otherwise indicate it as a 0.
```{r}
error = 0
for (i in 1:dim(week)[1]){
  #i
  sub = week [-(i:i),]
  glm.fit = glm(Direction ~ Lag1 + Lag2, data=sub, family=binomial)
  #ii
  glm.probs=predict(glm.fit, week[i:i,], type="response")
  #iii
  glm.pred=rep("Down",1)
  glm.pred[glm.probs>.5]="Up"
  #iv
  if(glm.pred!=week[i:i,]$Direction){
    error = error + 1
  }
}
```

#### 1e) Take the average of the n numbers obtained in (d)iv in order to obtain the LOOCV estimate for the test error. Comment on the results.
```{r}
error / 1089
```
Based on this analysis, I don't believe we should trust our model to make predictions. Considering the error rate is 45%, we should reassess the model or go into an entirely different direction.